Convert NSF data

First, work with the Carnegie classifications table to get the counts for individual institution types.
These are not complete, so won't sum up with the overall totals.

```{r}
source("functions.R")
carnegies = loadCarnegieTable()
table = cleanupTable(carnegies)
table = reformatDisciplines(table)
writeOutResults(table[table$level=="Bachelor's Degrees",])
```


Then, get the individual school counts. I'm just concatenating together a bunch of individual webcaspar school files for this.
``` {r}
schools = loadSchoolTable(c("rawdata/Schmidt.csv","rawdata/Schools2.csv"))
table = cleanupTable(schools)
table = reformatDisciplines(table)
writeOutResults(table[table$level=="Bachelor's Degrees",])
cat("schoolnames =",toJSON(c("",sort(unique(table$institution)))),"longdata",file="listOfSchools.json")
head(schools)
```

``` {r}
elitePublic = c("University of California-Berkeley","University of California-Los Angeles","University of Colorado at Boulder","University of Illinois at Urbana-Champaign","University of Iowa","University of Michigan at Ann Arbor","University of Minnesota - Twin Cities","University of Nebraska at Lincoln","University of Texas at Austin","University of Virginia, Main Campus","University of Washington - Seattle","University of Wisconsin-Madison")

Ivy = c("Harvard University","Princeton University","Yale University","Dartmouth College","University of Pennsylvania","Cornell University","Brown University","Columbia University")
new  = table[table$institution %in% Ivy | table$institution %in% elitePublic,]
new$institution[new$institution %in% Ivy] = "Ivy League"
new$institution[new$institution %in% elitePublic] = "Selected State Flagships"
table(new$institution)
writeOutResults(new[new$level=="Bachelor's Degrees",])
```


```{r}

table = read.table("DegreeData.csv",sep=",",skip = 6,header=T
           ,colClasses = c("character","factor","factor","character","character"),na.strings=c(".",'"."','"######"'))

names(table) = c("year","institution","discipline","gender","majors","minors")
#

if (FALSE) {
table = read.table("Schools.csv",sep=",",skip = 0,header=T
           ,colClasses = c("character","factor","factor","character","character"),na.strings=c(".",'"."','"######"'))
head(table)
names(table) = c("year","gender","degree","discipline","institution","muck","majors");
table=table[table$degree=="Bachelor's Degrees",]
table$minors = 0
}

table = cleanupTable(table)

table = reformatDisciplines(table)
head(table)
require(reshape2)
limited = melt(xtabs(majors~year+gender+field,table))
names(limited) = c("year","gender","field","count")
totals = ddply(limited,.(year),function(row){data.frame(total=sum(row$count))})
limited = merge(limited,totals)
head(limited)
write.table(limited,file="/Library/WebServer/Documents/Allfields/data.tsv",sep="\t",row.names=F)

ddply(table,.(institution),function(mytable){
  limited = melt(xtabs(majors~year+gender+field,mytable))
  cat(as.character(mytable$institution[1]))
  names(limited) = c("year","gender","field","count")
  totals = ddply(limited,.(year),function(row){data.frame(total=sum(row$count))})
  limited = merge(limited,totals)
  head(limited)
  write.table(limited,file=paste0("/Library/WebServer/Documents/Allfields/",gsub("/",".",mytable$institution[1]),".tsv"),sep="\t",row.names=F)
})

limited = melt(xtabs(majors~year+gender+field,table[table$institution=="Master's/Comprehensive II",]))

```

Long Term humanities degrees are stored in a separate table. 

```{r}
library(plyr,reshape2)
library(reshape2)
humanities = read.csv("FullHumanities.csv")
humanities = melt(humanities,id.vars="year")
humanities$gender = sapply(strsplit(as.character(humanities$variable),"\\."),'[[',1)
humanities$field = gsub(".*\\.","",humanities$variable)
table(humanities$field)
humanities = humanities[humanities$field %in% c("AreaStudies","Arts","Classics","English","History","Journalism","Languages","Music","Philosophy","Religion","total"),]
humanities = humanities[humanities$year < 1966,]
humanities$gender = factor(humanities$gender,levels = c("M","F"),labels=c("Male","Female"))
humanities = humanities[,-2]
names(humanities) = c("year","count","gender","field")
```

The older tables are yet messier, and missing a few years in the early 1960s.
```{r}
degrees = read.csv("OldHEGIS.csv")
mine = melt(degrees,id.vars=c("X1969.HEGIS.code","X1969.HEGIS.Descriptor","X1953.Order","X1952.Order","X1948.Order","Order64"))
table(mine$variable)
mine$variable = gsub("([MW]?)([A-Z][A-Z]+)(\\d\\d)","X\\3\\1\\2",mine$variable)
regex="X(\\d\\d)([MW]?)([A-Z][A-Z]+)"
mine$good = grepl(regex,mine$variable)
mine$Degree = gsub(regex,"\\3",as.character(mine$variable))
mine$Gender =  gsub(regex,"\\2",as.character(mine$variable))
mine$Year =  as.numeric(gsub(regex,"\\1",as.character(mine$variable))) + 1900
mine$field = mine$X1969.HEGIS.Descriptor
mine$HEGIS = mine$X1969.HEGIS.code
mine$value = as.numeric(mine$value)
mine = mine[,!grepl("^X",names(mine))]
mine = mine[mine$Gender %in% c("M","F","W"),]
generalCodes = mine[as.numeric(as.character(mine$HEGIS)) %% 100==0,c("field","HEGIS")]
generalCodes = generalCodes[!duplicated(generalCodes$HEGIS),]
head(mine)
mine = mine[!mine$field %in% c("History","Journalism","English","English, Literature, Creative Writing, Language Arts, Comparative Literature","Music, Sacred Music","Foreign Language and Literature","Religion","Spanish","French","German","Japanese","Chinese","Italian","Russian"),]
head(mine)
names(mine) = tolower(names(mine))
mine$gender = factor(mine$gender,levels = c("M","W"),labels=c("Male","Female"))
head(mine)
oldtotals = mine[mine$degree=="BA",c("year","value","gender","field")]
names(oldtotals) = c("year","count","gender","field")
```

Merging the two old series together:

```{r}
humanities$discipline=humanities$field
oldtotals$discipline = oldtotals$field
totals = ddply(humanities[humanities$field=="total",],.(year),function(frame) {
  return(data.frame(total=sum(frame$count)))
})
sums = rbind(oldtotals[!oldtotals$field=="Grand Total of All degrees",],humanities[!humanities$field=="total",])
head(sums)
sums$discipline = as.character(sums$discipline)
sums$field = as.character(sums$field)
require(plyr)
crosswalk = read.csv("lookups.csv",stringsAsFactors=F)
replacements = ddply(crosswalk,.(field),function(broadfield){
  data.frame(replacement = unique(sums$discipline[grep(paste(broadfield$regex,collapse="|"),sums$discipline)]))
},.progress="text")

sums$field[sums$field %in% replacements$replacement] = replacements$field[match(sums$field[sums$field %in% replacements$replacement],replacements$replacement)]
sums = ddply(sums,.(field,year,gender),function(field){
  data.frame(count=sum(field$count,na.rm=T))
})
counts = xtabs(count~field,sums)
sort(-counts[counts>0])
previousOutput = read.table("/Library/WebServer/Documents/Allfields/data.tsv",header=T,stringsAsFactors=F)

newtotals = rbind(totals,ddply(previousOutput,.(year), function(frame) {return(data.frame(total=frame$total[1]))}))
new = rbind(previousOutput,merge(sums,totals))

new[sample(1:nrow(new),10),]
new = new[new$field!="",]
library(reshape2)
better = melt(xtabs(count~year+field+gender,new))
names(better) = c("year","field","gender","count")
better = better[better$field %in% names(xtabs(count~field,better)[xtabs(count~field,better)>0]),]
better = merge(better,newtotals,by=c("year"))
write.table(better[order(better$year,better$field,better$gender),],file="/Library/WebServer/Documents/Allfields/longData.tsv",row.names=F,sep="\t")
```